Generalized Quasi-linear Approximation and Non-normality in Taylor-Couette Spiral Turbulence

### 2022-08-16 paper/paper.tex
Updates from Overleaf

%%%% Abstract text to be placed here %%%%%%%%%%%%
\begin{abstract}
Taylor-Couette flow is well known to admit a spiral turbulence state in which laminar and turbulent patches coexist around the cylinder. This flow state is quite complex, with delicate internal structure, and it can be traced into certain regimes of linear stability. This behavior is believed to be connected to the non-normality of the linear operator, which is itself a function of the control parameters. Using spiral turbulence in both linearly stable and unstable regimes, we investigate the effectiveness of the generalized quasilinear approximation (GQL), an extension of quasilinear theory designed to capture the essential aspects of turbulent flows. We find that GQL performs much better in the supercritical regime than the subcritical. By including only a small number of modes into the non-linear interactions, GQL simulations maintain a turbulent-like state when in the supercritical regime. However, a much larger number is required to avoid returning to the laminar state when in the subcritical regime. 
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Some of the most common forms of turbulence in nature are anisotropic, with a symmetry broken by rotation, magnetic fields, or boundary effects. In these situations, significant theoretical progress can be made by considering \emph{quasi-linear} (QL) models, in which one retains only those non-linearities that interact with the mean flow. Recently, the generalized quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively (see \cite{2022arXiv220505513M} for a recent review). GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows including forced zonal jet formation \cite{2016PhRvL.116u4501M}, rotating plane Couette flow \cite{2017JFM...810..412T}, rapidly rotating convection \cite{2018RSPSA.47480422T}, and channel flow\cite{2019Kellam}.

Here, we consider Taylor-Couette (TC) flows at moderate inner and outer Reynolds numbers, $|\Reyn_{i,o}| < 3500$. This is an interesting testbed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the cascade, local in coefficient space, found at high $\Reyn$. 
One of the key feature of TC flow is that it has non-linear and linear instabilities in different regions of parameter space. In the linearly stable regions, transient growth has been observed in TC flow\cite{2002PhFl...14.3475H, 2002PhFl...14.1655M}. 
We can write the Navier-Stokes equations in schematic form as
\begin{equation}
    \label{eqn:NS-schematic}
    \frac{\partial \mathbf{q}}{\partial t} + \Lop \mathbf{q} = \mathcal{N}(\mathbf{q}),
\end{equation}
where $\Lop$ is a linear operator and $\mathcal{N}$ is a non-linear function of the state vector $\mathbf{q} = \left[ \mathbf{u},  p\right]^T$, where $\mathbf{u}$ is the velocity field and $p$ the pressure.
As noted by \cite{2002PhFl...14.3475H}, the non-normality of the linear operator $\Lop$ for TC flow is a function of its control parameters.
That is, its commutator with its adjoint
\begin{equation}
    [\Lop^\dagger, \Lop] = f(\mu, \eta, \Reyn_i, \Gamma)
\end{equation}
can be zero or non-zero. 
This provides a very useful arena to test the validity of GQL in a system that admits a varying degree of non-normality.

As a first experiment to this end, we consider the \emph{spiral turbulence} flow regime. The classic paper by Coles \cite{1965JFM....21..385C} was the first to note a hysteretic transition to spiral turbulence when the cylinders were counter-rotating: an experiment at fixed $\Reyn_o$ will have spiral turbulence at a certain $\Reyn_i$ if $\Reyn_i$ is reached by \emph{decreasing} from a higher value, but will not have spiral turbulence at the same $\Reyn_i$ if it is reached by \emph{increasing} from a lower value. More recent work has identified both supercritical and subcritical paths to spiral turbulence\cite{2009PhRvE..80d6315M}. In that work, the authors follow spiral turbulence along a decreasing $\Reyn_i$ path and show that the state is maintained even below the linear stability boundary when $\Reyn_o = -3000$ but it relaminarizes below the stability boundary when $\Reyn_o = -1200$.
Using \cite{2009PhRvE..80d6315M} as a guide, we use direct numerical simulations (DNS) at $\Reyn_o = -3398$ and decrease $\Reyn_i$ into the subcritical region. We then run a suite of GQL simulations in both the super- and subcritical regions.

We begin in section~\ref{sec:methods} by detailing our equations, non-dimensionalization, and numerical methods. Section~\ref{sec:linear} reviews both modal and non-modal linear analyses of spiral turbulence. In section~\ref{sec:nonlinear}, we describe the main results of the paper, the comparison of DNS, QL, and GQL for supercritical and subcritical spiral turbulence. Finally, we offer discussion and concluding results in section~\ref{sec:conclusion}.
%Recently, a subcritical transition to turbulence via interpenetrating spirals has been reported \cite{2020JFM...892A..12C}. This work has a small aspect ratio $\Gamma = 5.26$, which delays the onset of linear instability to considerably higher $\Reyn_1$ for a given $\Reyn_2$.

@@

We non-dimensionalise lengths in terms of the gap width $d = R_2 - R_1$ and velocities in terms of the inner cylinder velocity $U_i = R_i \Omega_i$. Taylor-Couette flow can be parameterised in terms of four parameters, the inner cylinder Reynolds number $\Reyn_i$, the ratio of inner to outer radii $\eta \equiv R_i/R_o$, the ratio of outer to inner rotation rates $\mu \equiv \Omega_o/\Omega_i$, and the aspect ratio of the cylinder $\Gamma \equiv L_z/d$. In our computations, the $z$ direction is periodic and thus $L_z$ is the periodicity length. We will also refer to the two Reynolds numbers $\Reyn_i$, $\Reyn_o$, related by
\begin{equation}
    \frac{\Reyn_o}{\Reyn_i} = \frac{\mu}{\eta}.
\end{equation}
It is also quite useful to characterize times in terms of the viscous diffusion time $\tau_\nu \equiv d^2/\nu$. In our units, $\tau_\nu = \Reyn_i$.

@@

In order to demonstrate the effects non-normality plays, figure~\ref{fig:pseudospectra} shows both the modal spectrum (in blue points) and the $\epsilon-$pseudospectrum for the supercritical and subcritical $\Reyn_i$. The contour value $\epsilon^{-1}$ represents the set of values 
\begin{equation}
\label{eqn:pseudospectrum}
    \lVert (\sigma \mathbf{I} - \Lop)^{-1} \rVert \geq \epsilon^{-1}
\end{equation}
for a given complex number $\sigma = \gamma + i\omega$ here representing the eigenvalue corresponding to the temporal evolution of a linear solution, $e^{\sigma t}$ (see \cite{trefethen2005spectra} for an introduction). Equation~\ref{eqn:pseudospectrum} depends on the choice of norm; here we use the standard energy norm. These contours in turn give a measure of the maximum amplification of a linear disturbance: the further negative contours extend into the upper-half plane, the larger a linear disturbance can grow as a result of non-normality, irrespective of the presence of instability. This linear effect is captured entirely by GQL (and QL, for that matter). What role it plays in the development of subcritical spiral turbulence remains an open question.
\begin{figure}
    \centering
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/TC_pseudospectrum_Re1_900.00_eta_0.8830_mu_-3.3333_nr_64_m_6_kz_ 9.68.pdf}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/TC_pseudospectrum_Re1_700.00_eta_0.8830_mu_-4.2857_nr_64_m_6_kz_ 9.68.pdf}
    \end{minipage}
    \caption{Spectra (blue points) and pseudospectra (contours) for $\Reyn_i = 900$ (left) and $\Reyn_i = 700$ (right) for a mode with $m=6$ and $k_z = 9.68$. The latter is the z wavenumber at the critical $\Reyn_{i,crit} \simeq 778$. $\omega$ is the frequency and $\gamma$ the growth rate at each point in the domain. The stability boundary is marked with a gray line at $\gamma = 0$. The are seven equally spaced contours from $-1.8$ to 0.}
    \label{fig:pseudospectra}
\end{figure}
\section{Results}
\label{sec:nonlinear}
Figure~\ref{fig:urms_tz_rei700} shows a typical DNS snapshot of the rms velocity perturbation $u'_{rms}$ in the middle of the gap for $\Reyn_i=700$. Despite being well below the linear stability threshold $\Reyn_{i,crit} \simeq 778$, the flow morphology is nearly identical to the supercritical case $\Reyn_i = 900$ (not shown). 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{../figs/urms_tz_rei_700_reo-3398_000278.png}
    \caption{Slice of fluctuating rms velocity in the middle of the radial domain at $\Reyn_i = 700$, $\Reyn_o = -3398$. This solution is well below the linear stability threshold at }
    \label{fig:urms_tz_rei700}
\end{figure}

\subsection{Supercritical Spiral Turbulence}
We first consider the supercritical case with $\Reyn_i = 900$. We initialize GQL simulations with $\Lambda = 0, 1, 3, 5, 10, 30, 60$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.  Figure~\ref{fig:rei900_snapshots} shows slices at mid-gap, $r = R_{mid}$. The first important thing to note is that low-order models are not able to maintain the spiral structure until $\Lambda \geq 30$. However, even including a single additional low mode, $\Lambda = 1$, shows a significant increase in flow complexity compared to QL. It is quite interesting to note that QL \emph{does} feature some degree of spatial intermittency--there are clearly separated regions hosting spiral waves interspersed with featureless, laminar patches. 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figs/rei900_snapshots.pdf}
    \caption{RMS velocity perturbations in the $z-\theta$ plane at $r=R_{mid}$ for $\Reyn_i=900, \Reyn_0=-3398$. The upper left panel shows DNS, the upper right panel shows QL, and the remaining four are GQL with $\Lambda = 1,3,5,10$. Until $\Lambda > 30$, spiral turbulence is not apparent.}
    \label{fig:rei900_snapshots}
\end{figure}

Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs. 
The supercritical case shows significant oscillations at all $\Lambda < 30$.
While both $\Lambda = 30$ and 60 show spiral turbulence, $\Lambda=30$ has considerably more wave motion in the laminar region than either $\Lambda = 60$ or DNS. 
Interestingly, figure~\ref{fig:ke_vs_t_rei900} shows that $\Lambda=30$ has a \emph{lower} kinetic energy much closer to that of DNS than does $\Lambda=60$.
By examining power spectra in figure~\ref{fig:rei900_spectra}, we begin to see the reason for the more complex flows seen with $\Lambda > 0$. Even low-order non-linear scatter leads to a steady increase 
\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei700_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$. $\Reyn_i = 900$ (left), started from small perturbations, is above the linear stability boundary. $\Reyn_i = 700$ (right panel), continued by lower $\Reyn_i$ from previous simulations, is in the hysteretic region below the stability boundary. Starting from a saturated spiral turbulence state from DNS at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10,30,60$. At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations. At lower $\Reyn_i$, $\Lambda = 60$ is required to maintain the spiral turbulence state. }
    \label{fig:ke_vs_t_rei900}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{../figs/rei900_spectra.pdf}
    \caption{$\theta$-$z$ power spectra of $v$ at midgap for $\Reyn_i = 900$ at various GQL cutoffs. Adding a single GQL mode significantly improves the power spectrum. Spiral turbulence appears at $\Lambda=30$. The obvious square in $\Lambda = 60$ represents a sudden drop in spectral power at the cutoff wavenumber.}
    \label{fig:rei900_spectra}
\end{figure}
Focusing on $\Lambda=5$, figure~\ref{fig:rei900_lambda5_story} shows that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns originally identified in exhaustive series of experiments reported in \cite{1984JFM...146...45M} and subsequently found in simulations (e.g. \cite{2009PhRvE..80d6315M}). At the time of kinetic energy minimum, labelled 1 in the left hand side of figure~\ref{fig:rei900_lambda5_story}, the flow seems to recapitulate the interpenetrating spirals (IPS) characteristic of lower $\Reyn$ counter-rotating. As the kinetic energy rises, reaching a knee at point 2, the flow morphology resembles the patchy bursts characteristic of the intermittency regime (INT). Finally, at the peak of the kinetic energy oscillation, point 3 in figure~\ref{fig:rei900_lambda5_story}, the flow reaches its closest approach to something resembling spiral turbulence, though it is very difficult to identify the flow as such from a slice at a single point in time and plane in space.

@@

\section{Conclusion}
\label{sec:conclusion}
We have presented a series of simulations demonstrating the performance of GQL at various cutoffs for Taylor-Couette flow in the spiral turbulence regime. GQL is a significant improvement over QL models in representing statistical quantities, particularly the power spectra. However, it does not retain an unambiguous spiral form in the supercritical regime until $\Lambda \simeq 30$; at subcritical $\Reyn_i$, $\Lambda \simeq 60$ is required to see any sustained flows at all. Interestingly, when GQL solutions do sustain non-CCF flows in the subcritical regime, they also produce spiral patterns. Prior work has focused on the utility of GQL in presenting a more accurate \emph{statistical} picture of turbulent flows than QL \cite{2016PhRvL.116u4501M, 2017JFM...810..412T,2018RSPSA.47480422T,2019Kellam}. This focus is particularly important in light of its potential as as means of improving the accuracy of direct statistical simulation (DSS), a method for studying turbulence by computing its statistics directly, rather than individual realizations \cite{2022arXiv220505513M}. 
We have shown here that it is possible to use GQL as a means to investigate the non-linear interactions of greatest importance for maintaining the intricate patterns manifested in Taylor-Couette flow. In future work, we will consider the ability of GQL to maintain exact coherent states identified by Deguchi et al \cite{2014PhRvL.112r4502D} and the dynamics of which were studied recently by Wang and coworkers \cite{2022arXiv220712990W}. Understanding how these interactions work to shape subcritical TC turbulence represents a promising new avenue to understanding how ECS are maintained and how they in turn sustain turbulence below linear stability bounds. In particular, one can begin to understand the relative importance of non-normal effects (represented entirely with GQL) and non-linear ECS.

>>>>

%%%% Abstract text to be placed here %%%%%%%%%%%%
\begin{abstract}
Taylor-Couette flow is well known to admit a spiral turbulence state in which laminar and turbulent patches coexist around the cylinder. This flow state is quite complex, with delicate internal structure, and it can be traced into certain regimes of linear stability. This behaviour is believed to be connected to the non-normality of the linear operator, which is itself a function of the control parameters. Using spiral turbulence in both linearly stable and unstable regimes, we investigate the effectiveness of the generalised quasilinear approximation (GQL), an extension of quasilinear theory designed to capture the essential aspects of turbulent flows. We find that GQL performs much better in the supercritical regime than the subcritical. By including only a small number of modes into the non-linear interactions, GQL simulations maintain a turbulent-like state when in the supercritical regime. However, a much larger number is required to avoid returning to the laminar state when in the subcritical regime. 
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Some of the most common forms of turbulence in nature are anisotropic, with a symmetry broken by rotation, magnetic fields, or boundary effects. In these situations, significant theoretical progress can be made by considering \emph{quasi-linear} (QL) models, in which one retains only those non-linearities that interact with the mean flow. Recently, the generalised quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively (see \cite{2022arXiv220505513M} for a recent review). GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows including forced zonal jet formation \cite{2016PhRvL.116u4501M}, rotating plane Couette flow \cite{2017JFM...810..412T}, rapidly rotating convection \cite{2018RSPSA.47480422T}, and channel flow\cite{2019Kellam}.

Here, we consider Taylor-Couette (TC) flows at moderate inner and outer Reynolds numbers, $|\Reyn_{i,o}| < 3500$. This is an interesting test bed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the cascade, local in coefficient space, found at high $\Reyn$. 
One of the key feature of TC flow is that it has non-linear and linear instabilities in different regions of parameter space. In the linearly stable regions, transient growth has been observed in TC flow\cite{2002PhFl...14.3475H, 2002PhFl...14.1655M}. 
We can write the Navier-Stokes equations in schematic form as
\begin{equation}
    \label{eqn:NS-schematic}
    \frac{\partial \mathbf{q}}{\partial t} + \Lop \mathbf{q} = \mathcal{N}(\mathbf{q}),
\end{equation}
where $\Lop$ is a linear operator and $\mathcal{N}$ is a non-linear function of the state vector $\mathbf{q} = \left[ \mathbf{u},  p\right]^T$, where $\mathbf{u}$ is the velocity field and $p$ the pressure.
As noted by \cite{2002PhFl...14.3475H}, the non-normality of the linear operator $\Lop$ for TC flow is a function of its control parameters.
That is, its commutator with its adjoint
\begin{equation}
    [\Lop^\dagger, \Lop] = f(\mu, \eta, \Reyn_i, \Gamma)
\end{equation}
can be zero or non-zero. 
This provides a very useful arena to test the validity of GQL in a system that admits a varying degree of non-normality.

As a first experiment to this end, we consider the \emph{spiral turbulence} flow regime. The classic paper by Coles \cite{1965JFM....21..385C} was the first to note a hysteretic transition to spiral turbulence when the cylinders were counter-rotating: an experiment at fixed $\Reyn_o$ will have spiral turbulence at a certain $\Reyn_i$ if $\Reyn_i$ is reached by \emph{decreasing} from a higher value, but will not have spiral turbulence at the same $\Reyn_i$ if it is reached by \emph{increasing} from a lower value. More recent work has identified both supercritical and subcritical paths to spiral turbulence\cite{2009PhRvE..80d6315M}. In that work, the authors follow spiral turbulence along a decreasing $\Reyn_i$ path and show that the state is maintained even below the linear stability boundary when $\Reyn_o = -3000$ but it relaminarises below the stability boundary when $\Reyn_o = -1200$.
Using \cite{2009PhRvE..80d6315M} as a guide, we use direct numerical simulations (DNS) at $\Reyn_o = -3398$ and decrease $\Reyn_i$ into the subcritical region. We then run a suite of GQL simulations in both the super- and subcritical regions.

We begin in section~\ref{sec:methods} by detailing our equations, non-dimensionalisation, and numerical methods. Section~\ref{sec:linear} reviews both modal and non-modal linear analyses of spiral turbulence. In section~\ref{sec:nonlinear}, we describe the main results of the paper, the comparison of DNS, QL, and GQL for supercritical and subcritical spiral turbulence. Finally, we offer discussion and concluding results in section~\ref{sec:conclusion}.
%Recently, a subcritical transition to turbulence via interpenetrating spirals has been reported \cite{2020JFM...892A..12C}. This work has a small aspect ratio $\Gamma = 5.26$, which delays the onset of linear instability to considerably higher $\Reyn_1$ for a given $\Reyn_2$.

@@

We non-dimensionalise lengths in terms of the gap width $d = R_2 - R_1$ and velocities in terms of the inner cylinder velocity $U_i = R_i \Omega_i$. Taylor-Couette flow can be parameterised in terms of four parameters, the inner cylinder Reynolds number $\Reyn_i$, the ratio of inner to outer radii $\eta \equiv R_i/R_o$, the ratio of outer to inner rotation rates $\mu \equiv \Omega_o/\Omega_i$, and the aspect ratio of the cylinder $\Gamma \equiv L_z/d$. In our computations, the $z$ direction is periodic and thus $L_z$ is the periodicity length. We will also refer to the two Reynolds numbers $\Reyn_i$, $\Reyn_o$, related by
\begin{equation}
    \frac{\Reyn_o}{\Reyn_i} = \frac{\mu}{\eta}.
\end{equation}
It is also quite useful to characterise times in terms of the viscous diffusion time $\tau_\nu \equiv d^2/\nu$. In our units, $\tau_\nu = \Reyn_i$.

@@

In order to demonstrate the effects non-normality plays, figure~\ref{fig:pseudospectra} shows both the modal spectrum (in blue points) and the $\epsilon-$pseudospectrum for the supercritical and subcritical $\Reyn_i$. The contour value $\epsilon^{-1}$ represents the set of values 
\begin{equation}
\label{eqn:pseudospectrum}
    \lVert (\sigma \mathbf{I} - \Lop)^{-1} \rVert \geq \epsilon^{-1}
\end{equation}
for a given complex number $\sigma = \gamma + i\omega$ here representing the eigenvalue corresponding to the temporal evolution of a linear solution, $e^{\sigma t}$ (see \cite{trefethen2005spectra} for an introduction). Equation~\ref{eqn:pseudospectrum} depends on the choice of norm; here we use the standard energy norm. These contours in turn give a measure of the maximum amplification of a linear disturbance: the further negative contours extend into the upper-half plane, the larger a linear disturbance can grow as a result of non-normality, irrespective of the presence of instability. This linear effect is captured entirely by GQL (and QL, for that matter). What role it plays in the development of subcritical spiral turbulence remains an open question, though our results provide some interesting constraints.
\begin{figure}
    \centering
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/TC_pseudospectrum_Re1_900.00_eta_0.8830_mu_-3.3333_nr_64_m_6_kz_ 9.68.pdf}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/TC_pseudospectrum_Re1_700.00_eta_0.8830_mu_-4.2857_nr_64_m_6_kz_ 9.68.pdf}
    \end{minipage}
    \caption{Spectra (blue points) and pseudospectra (contours) for $\Reyn_i = 900$ (left) and $\Reyn_i = 700$ (right) for a mode with $m=6$ and $k_z = 9.68$. The latter is the z wavenumber at the critical $\Reyn_{i,crit} \simeq 778$. $\omega$ is the frequency and $\gamma$ the growth rate at each point in the domain. The stability boundary is marked with a grey line at $\gamma = 0$. The are seven equally spaced contours from $-1.8$ to 0.}
    \label{fig:pseudospectra}
\end{figure}
\section{Results}
\label{sec:nonlinear}
Figure~\ref{fig:urms_tz_rei700} shows a typical DNS snapshot of the RMS velocity perturbation $u'_{rms}$ in the middle of the gap for $\Reyn_i=700$. Despite being well below the linear stability threshold $\Reyn_{i,crit} \simeq 778$, the flow morphology is nearly identical to the supercritical case $\Reyn_i = 900$ (not shown). 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{../figs/urms_tz_rei_700_reo-3398_000278.png}
    \caption{Slice of fluctuating RMS velocity in the middle of the radial domain of a DNS at $\Reyn_i = 700$, $\Reyn_o = -3398$. This solution is well below the linear stability threshold at $\Reyn_{i,crit} \simeq 778$.}
    \label{fig:urms_tz_rei700}
\end{figure}

\subsection{Supercritical Spiral Turbulence}
We first consider the supercritical case with $\Reyn_i = 900$. We initialise GQL simulations with $\Lambda = 0, 1, 3, 5, 10, 30, 60$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.  Figure~\ref{fig:rei900_snapshots} shows slices at mid-gap, $r = R_{mid}$. The first important thing to note is that low-order models are not able to maintain the spiral structure until $\Lambda \geq 30$. However, even including a single additional low mode, $\Lambda = 1$, shows a significant increase in flow complexity compared to QL. It is quite interesting to note that QL \emph{does} feature some degree of spatial intermittency--there are clearly separated regions hosting spiral waves interspersed with featureless, laminar patches. 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figs/rei900_snapshots.pdf}
    \caption{RMS velocity perturbations in the $z-\theta$ plane at $r=R_{mid}$ for $\Reyn_i=900, \Reyn_0=-3398$. The upper left panel shows DNS, the upper right panel shows QL, and the remaining six are GQL with $\Lambda = 1,3,5,10, 30, 60$. Until $\Lambda > 30$, spiral turbulence is not apparent.}
    \label{fig:rei900_snapshots}
\end{figure}

Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs. 
The supercritical case shows significant oscillations at all $\Lambda < 30$.
While both $\Lambda = 30$ and 60 show spiral turbulence, $\Lambda=30$ has considerably more wave motion in the laminar region than either $\Lambda = 60$ or DNS. 
Interestingly, figure~\ref{fig:ke_vs_t_rei900} shows that $\Lambda=30$ has a \emph{lower} kinetic energy much closer to that of DNS than does $\Lambda=60$.

By examining power spectra in figure~\ref{fig:rei900_spectra}, we begin to see the reason for the more complex flows seen with $\Lambda > 0$. Even low-order non-linear scatter leads to a steady increase to higher $k$ and $m$. The spectra of $\Lambda = 30$ and 60 are of particular interest in understanding the nature of spiral turbulence: they show a transition toward filamentary structure aligned along the diagonal, very similar to DNS. The square feature clearly visible at $\Lambda = 60$ is analogous to similar structures seen in channel flow turbulence (see figure 4.20 of \cite{2019Kellam}). In that work, the author proposed a potential explanation in terms of the linear instability of the non-linear solution: assume there is an instability of the mean $(m, k_z) = (0,0)$ at some wavenumber $k'$. If $k'$ is counted among the high modes, high-high $\to$ high interactions smoothly populate the high modes. However, as $\Lambda$ increases, it reaches a point at which the last unstable mode $k'$ becomes counted among the low modes. Because low-low $\to$ high interactions are truncated in GQL,  this results in a sudden decrease in the amount of energy available in the high part of the spectrum. Reference \cite{2019Kellam} constructs a preliminary model of this behaviour for plane channel flow and finds support for the idea.
\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei700_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$. $\Reyn_i = 900$ (left), started from small perturbations, is above the linear stability boundary. $\Reyn_i = 700$ (right panel), continued by lower $\Reyn_i$ from previous simulations, is in the hysteretic region below the stability boundary. Starting from a saturated spiral turbulence state from DNS at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10,30,60$. At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations. At lower $\Reyn_i$, $\Lambda = 60$ is required to maintain the spiral turbulence state. }
    \label{fig:ke_vs_t_rei900}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{../figs/rei900_spectra.pdf}
    \caption{$\theta$-$z$ power spectra of $v$ at mid-gap for $\Reyn_i = 900$ at various GQL cutoffs. Adding a single GQL mode significantly improves the power spectrum. Spiral turbulence appears at $\Lambda=30$. The obvious square in $\Lambda = 60$ represents a sudden drop in spectral power at the cutoff wavenumber. The figure shows only the lowest 100 modes in each direction, though in the simulations $k_{max} = m_{max} = 512$. }
    \label{fig:rei900_spectra}
\end{figure}
Focusing on $\Lambda=5$, figure~\ref{fig:rei900_lambda5_story} shows that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns originally identified in exhaustive series of experiments reported in \cite{1984JFM...146...45M} and subsequently found in simulations (e.g. \cite{2009PhRvE..80d6315M}). At the time of kinetic energy minimum, labelled 1 in the left hand side of figure~\ref{fig:rei900_lambda5_story}, the flow seems to recapitulate the interpenetrating spirals (IPS) characteristic of lower $\Reyn$ counter-rotating. As the kinetic energy rises, reaching a knee at point 2, the flow morphology resembles the patchy bursts characteristic of the intermittency regime (INT). Finally, at the peak of the kinetic energy oscillation, point 3 in figure~\ref{fig:rei900_lambda5_story}, the flow reaches its closest approach to something resembling spiral turbulence, though it is very difficult to identify the flow as such from a slice at a single point in time and plane in space.

@@

\section{Conclusion}
\label{sec:conclusion}
We have presented a series of simulations demonstrating the performance of GQL at various cutoffs for Taylor-Couette flow in the spiral turbulence regime. GQL is a significant improvement over QL models in representing statistical quantities, particularly the power spectra. However, it does not retain an unambiguous spiral form in the supercritical regime until $\Lambda \simeq 30$; at subcritical $\Reyn_i$, $\Lambda \simeq 60$ is required to see any sustained flows at all. Interestingly, when GQL solutions do sustain non-CCF flows in the subcritical regime, they also produce spiral patterns. Prior work has focused on the utility of GQL in presenting a more accurate \emph{statistical} picture of turbulent flows than QL \cite{2016PhRvL.116u4501M, 2017JFM...810..412T,2018RSPSA.47480422T,2019Kellam}. This focus is particularly important in light of its potential as as means of improving the accuracy of direct statistical simulation (DSS), a method for studying turbulence by computing its statistics directly, rather than individual realisations \cite{2022arXiv220505513M}. 
We have shown here that it is possible to use GQL as a means to investigate the non-linear interactions of greatest importance for maintaining the intricate patterns manifested in Taylor-Couette flow. In future work, we will consider the ability of GQL to maintain exact coherent states identified by Deguchi et al \cite{2014PhRvL.112r4502D} and the dynamics of which were studied recently by Wang and coworkers \cite{2022arXiv220712990W}. Understanding how these interactions work to shape subcritical TC turbulence represents a promising new avenue to understanding how ECS are maintained and how they in turn sustain turbulence below linear stability bounds. In particular, one can begin to understand the relative importance of non-normal effects (represented entirely with GQL) and non-linear ECS.
